---
title: "Multiple-Variable Linear Regression"
output:
  html_document:
    df_print: paged
---

R provides comprehensive support for multiple linear regression.

We shall consider the dataset "Regression.csv", which includes historical data on college enrollment numbers. Rarely in the real world is there ever a correlation between only two variables. Most of the time, several variables interact with one another in a predictable manner.

So, let's first import the dataset:
```{r}
regression <- read.csv("Regression.csv")
```
Now, we will use the lm() command in R to evaluate the linear regression of multiple variables (unemployment rate, high school graduates, and income) and the plot() command to provide checks for heteroscedasticity, normality, and influential observerations with diagnostic plots:
```{r}
fit <- lm(regression$ROLL ~ regression$UNEM + regression$HGRAD + regression$INC, data=regression)
fit
layout(matrix(c(1,2,3,4),2,2))
linearmodel = plot(fit)
linearmodel
```
It is clear what the regression equation is, and now we will predict the enrollment if the
Unemployment Rate = 4.1%
High School Graduates = 17,231
Income = $3,000
(Be sure to call up the variable "Enrollment" to print out the results!)
```{r}
UnemploymentRate = 4.1
HighSchoolGraduates = 17231
Income = 3000

Enrollment = -9153.2545 + UnemploymentRate * 450.1245 + HighSchoolGraduates * 0.4065 + Income * 4.2749

Enrollment
```
Now, we will use more plot functions to compare linear models to include or exclude one of the three influencing variables:
```{r}
fit1 <- lm(regression$ROLL ~ regression$UNEM + regression$HGRAD + regression$INC, data=regression)
fit2 <- lm(regression$ROLL ~ regression$UNEM + regression$HGRAD)
fit3 <- lm(regression$ROLL ~ regression$HGRAD + regression$INC)
fit4 <- lm(regression$ROLL ~ regression$UNEM + regression$INC)
anova(fit1, fit2, fit3, fit4)
```
Now, the analysis of the data is up to you!